Rates and delay times of type Ia supernovae in the Dark Energy Survey

### 2021-02-05 draft.tex

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Rates of SNe Ia in DES]{Measuring the Type Ia supernova delay time distribution via host galaxy rates in the Dark Energy Survey}

@@

% Abstract of the paper
\begin{abstract}
Measuring the rate of type Ia supernovae (SNe Ia) is crucial method by which to understand the progenitors and physical processes involved in these explosions. Since SNe Ia are a key source of heavy elements, the SN Ia rate is also an important ingredient in models of galaxy evolution. We use a sample of 1411 photometrically classified SNe discovered by the Dark Energy Survey (DES) along with \textbf{final number} field galaxies to calculate the rate of SNe Ia per galaxy as a function of galaxy stellar mass across a broad range of scales $8.5 \leq \log(M_*/\mathrm{M}_{\odot}) \leq 11.25$. We find that the SN Ia rate increases with stellar mass with a slope of $0.70 \pm 0.02$, which is consistent with previous work. We use an empirical model of stellar mass assembly to estimate the average star-formation histories (SFHs) of galaxies across the stellar mass range of our measurement. We combine the modelled SFHs with the SN Ia rates to estimate constraints on the SN Ia delay time distribution (DTD), and find a power-law with index $\beta = -0.96 \pm 0.5$ and normalisation $A = 1.7 \pm0.05 \times 10^{-13}~\mathrm{SNe}~{\mathrm{M}_{\odot}}^{-1}~\mathrm{yr}^{-1}$. 
\end{abstract}

@@

To correct for these incompleteness, we follow a similar method to that used in the PTF rates analyses of \citet{Frohmaier2019,Frohmaier2020} as well as the DES-SN superluminous SN rate of \citet{Thomas2020}. We run a suite of simulations using \texttt{SNANA} \citet{Kessler2009a}. We simulate 230,000 SNe in the redshift range $0.05 \leq z_{\mathrm{SN}} \leq 1.3$, and draw peak magnitudes from \textbf{ADD DETAILS}. We then run mock versions of the DES-SN survey, using the exact cadence, conditions, and zeropoints from the survey itself. All of the detected simulated SNe are passed through the light curve fit of SALT2 \citep{Betoule2014}, as per the implementation in \texttt{SNANA}, and those that fail the light curve cuts outlined in Section \ref{subsubsec:sn_classify} are discarded. We are left with a fraction of the original simulated SNe, and that fraction is dependent on location, explosion epoch, and redshift, which we define as the SN efficiency. For the $i$th SN the efficiency $\eta_{\mathrm{SN}, i}$ in field $F$, exploding at time $t_0$ at redshift $z$, is:
\begin{equation}
    \eta_{\mathrm{SN},i} (F_i,z_i,t_{0,i}) = \left( \frac{N_{\mathrm{obs}}\left(F_i,z_i,t_{0,i}\right)}{N_{\mathrm{sim}}\left(F_i,z_i,t_{0,i}\right)}\right)\,.
\end{equation}

>>>>

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Rates of SNe Ia in DES]{Rates and delay times of type Ia supernovae in the Dark Energy Survey}

@@

% Abstract of the paper
\begin{abstract}
We use a sample of 881 photometrically classified SNe discovered by the Dark Energy Survey (DES) in the redshift range $0.2 < z <0.6$, along with 40415 field galaxies to calculate the rate of type Ia supernovae (SNe Ia) per galaxy. We recover the established positive trend of SN Ia rate as a function of galaxy stellar mass across a broad range of scales $8.5 \leq \log(M_*/\mathrm{M}_{\odot}) \leq 11.25$. We find that the SN Ia rate increases with stellar mass roughly linearly in log space, with a slope of $0.63 \pm 0.02$, which is consistent with previous work. We use an empirical model of stellar mass assembly to estimate the average star-formation histories (SFHs) of galaxies across the stellar mass range of our measurement. Combining the modelled SFHs with the SN Ia rates to estimate constraints on the SN Ia delay time distribution (DTD), we find the data are best fit by a power-law DTD with index $\beta = -1.14 \pm 0.05$ and normalisation $A = 2.1 \pm0.05 \times 10^{-13}~\mathrm{SNe}~{\mathrm{M}_{\odot}}^{-1}~\mathrm{yr}^{-1}$. Upon splitting the SN sample by properties of the light curves, we find a strong dependence on DTD slope with the SN stretch parameter, with slower-declining SNe exhibiting a steeper DTD slope.
\end{abstract}

@@

To correct for these incompleteness, we follow a similar method to that used in the PTF rates analyses of \citet{Frohmaier2019,Frohmaier2020} as well as the DES-SN superluminous SN rate of \citet{Thomas2020}. We run a suite of simulations using an identical method to that described in detail in \citet{Vincenzi2020}. We simulate SNe using the SuperNova ANAlysis software (\texttt{SNANA} \citealt{Kessler2009a}) integrated into the \texttt{pippin} framework \citep{Hinton2020}. We simulate 230,000 SNe in the redshift range $0.05 \leq z_{\mathrm{SN}} \leq 1.3$. The SNe are generated with SEDs based upon the SALT2 model \citep{Guy2007} and trained on the Joint Lightcurve Analysis dataset \citep{Betoule2014}. The SALT2 parameters $x_1$ (stretch) and $c$ (colour) are drawn randomly from the intrinsic asymmetric Gaussian distributions described in \citet{Scolnic2016} and "blurred" following the intrinsic scatter model of \citet{Guy2010}, while redshifts are drawn following the volumetric rate evolution of \citet{Frohmaier2019}. The SNe are simulated with explosion epochs $t_0$ uniformly distributed between two months before DES-SN began and two months after it finished. We run mock versions of the DES-SN survey, using the exact cadence, conditions, and zeropoints from the survey itself. All of the detected simulated SNe are passed through the light curve fit of SALT2 \citep{Betoule2014}, as per the implementation in \texttt{SNANA}, and those that fail the light curve cuts outlined in Section \ref{subsubsec:sn_classify} are discarded. We are left with a fraction of the original simulated SNe, and that fraction is dependent on location, explosion epoch, and redshift, which we define as the SN efficiency. For the $i$th SN the efficiency $\eta_{\mathrm{SN}, i}$ in field $F$, exploding at time $t_0$ at redshift $z$, is:
\begin{equation}
    \eta_{\mathrm{SN},i} (F_i,z_i,t_{0,i}) = \left( \frac{N_{\mathrm{obs}}\left(F_i,z_i,t_{0,i}\right)}{N_{\mathrm{sim}}\left(F_i,z_i,t_{0,i}\right)}\right)\,.
\end{equation}
